### 2016-05-15 Manuscript/SW_early_life_telomeres.Rmd
Home edits

Because we observed apparent telomere lenghtening in our data, we used the entire Seychelles warbler dataset to test whether this lengthening could be explained by measurement error. We used all birds (adults and juveniles) with at least two telomere measurements (N = `r nrow(subset(ddL,Group == 'Among samples'))` samples from `r length(unique(subset(ddL,Group == 'Among samples')$BirdID))` birds). We then calculated $\Delta$RTL for each pair of samples (for example, for a bird with three logitudinal measurements throughout their life we calculated the change in RTL between the first and second sample and between the second and third sample). We then compared these longitudinal changes in telomere length to sampling error, using all samples where we had at least two repeat telomere measurements for the same sample (N = `r nrow(subset(ddL,Group == 'Within samples'))` measurements from `r length(unique(subset(ddL,Group == 'Within samples')$BirdID))` samples).

@@

To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we used the entire Seychelles warbler telomere dataset to compare variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance among individuals compared to among samples (Levene's test XXX). Importantly, splitting the longitudinal data revealed that not only did we observe significantly more shortening within individuals compared to within samples (XXX), but also significantly more lengthening (XXX).

>>>>

Because we observed apparent telomere lenghtening in our data, we used the entire Seychelles warbler dataset to test whether this lengthening could be explained by measurement error. We used all birds (adults and juveniles) with at least two telomere measurements (N = `r nrow(subset(ddL,Group == 'Among samples'))` samples from `r length(unique(subset(ddL,Group == 'Among samples')$BirdID))` birds). We then calculated $\Delta$RTL for each pair of samples (for example, for a bird with three logitudinal measurements throughout their life we calculated the change in RTL between the first and second sample and between the second and third sample). We then compared these longitudinal changes in telomere length to sampling error, using all samples where we had at least two repeat telomere measurements for the same sample (N = `r nrow(subset(ddL,Group == 'Within samples'))` measurements from `r length(unique(subset(ddL,Group == 'Within samples')$BirdID))` samples). We calculated $\Delta$RTL between pairs repeat measurements within samples in exactly the same way as for across samples, except that repeat measurements were ordered by the date at which samples were measured rather thjan by age.

@@

To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we used the entire Seychelles warbler telomere dataset to compare variance in telomere length among repeat measurements of samples to the variance observed among different samples of the same individual (see methods for details). We found significantly higher variance among individuals compared to among samples (Levene's test: F = `r round2(leveneTest(DeltaRTL~Group,data=ddL)['F value'][1,1])`; P `r round3(leveneTest(DeltaRTL~Group,data=ddL)['Pr(>F)'][1,1])`; Fig. 4C). Importantly, splitting the longitudinal data into instances of shortening ($\Delta$RTL < 0) and lenghtning ($\Delta$RTL > 0) revealed that not only did we observe significantly more shortening within individuals compared to within samples (Wilcoxon test: W = `r round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$statistic,0)`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL<0))$p.value)`), but also significantly more lengthening (W = `r as.character(round(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$statistic,0))`; P  `r round3(wilcox.test(DeltaRTL~Group,data=subset(ddL,DeltaRTL>0))$p.value)`).

### 2016-05-15 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
Home edits

Fig4B <- ggplot(temp,aes(x = Agemonths,y = CqGAPDH,group = BirdID,col=factor(BirdID)))+
geom_point()+
geom_line()+
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')
  annotate('text',x = -5,y = 2,label = 'B')

@@

ggplot(ddL,aes(y = DeltaTL,x = DeltaGAP, col = Group))+
geom_point()+
stat_smooth(method = 'lm')

plot(DeltaTL~Agemonths,data=dd3)
plot(DeltaGAP~Agemonths,data=dd3)

plot(DeltaGAP~TimeDiff,data=dd3)

hist(dd$CqGAPDH)

plot(CqTelomere~CqGAPDH,data=dd)

boxplot(CqTelomere~Ageclass,data=dd)
boxplot(CqGAPDH~Ageclass,data=dd)

ggplot(dd,aes(y = CqTelomere, x = CqGAPDH,col = Ageclass))+
geom_point()+
stat_smooth(method = 'lm')

>>>>

Fig4B <- ggplot(temp,aes(x = Agemonths,y = RTL,group = BirdID,col=factor(BirdID)))+
geom_point()+
geom_line()+
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')
  annotate('text',x = -5,y = 2,label = 'B')

@@
